---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# absorbedDose

<!-- badges: start -->
<!-- badges: end -->

The package  absorbedDose calculates the estimated organ absorbed dose
        following the Chernobyl Nuclear Power Plant accident.
        
The details about the model can be found in 
`r citation("absorbedDose")$textVersion`



## Installation

The package <code>absorbedDose</code> can be installed in the following way:

``` R
# install.packages("devtools")
devtools::install_github("absorbedDose/absorbedDose")
```

and then call

```{r libCall}
library("absorbedDose")
```

## Data set

A data set to calculate the doses should look like the generated sample data set <code>dose_data</code>: 

```{r data}
print(dose_data, nrow=10)
```

## Calculation of the absorbed dose

The function <code>calculate_dose</code> calls functions for data management 
(<code>dose_dm</code>, <code>dose_new_variables</code>),
for calculation of each dose component (<code>dose_external</code>, <code>dose_internal_ing</code>,
<code>dose_internal_milk</code>, <code>dose_inhalation</code>), 
and for adding the dose components <code>dose_total</code>). 
The doses are calculated per row (per year of follow-up time).

```{r example}
doses_row <- calculate_dose(dose_data, organ_ls=c("Colon", "Thyroid"))
```

To get the total absorbed dose under the follow-up period for each person 
(for each id) call:
```{r example2}
doses_id <- dose_total_per_person(doses_row)
```

