---
title: "absorbedDose"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{absorbedDose}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The package <tt>absorbedDose</tt> is developed to calculate the estimated organ absorbed dose
        following the Chernobyl Nuclear Power Plant accident.
The details about the model used for dose calculations can be found in 

`r citation("absorbedDose")$textVersion`.

```{r setup}
library(absorbedDose)
```

## Data set

A data set to calculate the doses should look like the generated sample data set <tt>dose_data</tt>. Each person has a unique id. 

```{r data}
head(dose_data)
```

## Example

The function <tt>calculate_dose()</tt> calls functions for data management 
(<tt>dose_dm()</tt>, <tt>dose_new_variables()</tt>),
for each dose component (<tt>dose_external()</tt>, <tt>dose_internal_ing()</tt>,
<tt>dose_internal_milk()</tt>, <tt>dose_inhalation()</tt>) and for the sum of the components (<tt>dose_total()</tt>). 
The doses are calculated per raw (per each year of the follow-up time).

```{r example}
doses <- calculate_dose(dose_data, organ_ls = c("Thyroid", "Colon"))
head(doses)
```

The function <tt>dose_total_per_person()</tt> sums the above calculated absorbed doses for the follow-up period for each person 
(for each id) and for each organ.
```{r example2}
doses_tot <- dose_total_per_person(doses)
head(doses_tot)
```
