---
title: "absorbedDose"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{absorbedDose}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The package <tt>absorbedDose</tt> is developed to calculate the estimated organ absorbed dose
        following the Chernobyl Nuclear Power Plant accident.
The details about the used model for dose calculation can be found in 

`r citation("absorbedDose")$textVersion`

The details about the model to calculate the number of attributable cancer cases and the effective dose
can be found in

Tondel M, Gabrysch K, Rääf C, Isaksson M. (2025)
A model for estimating radiation doses and population cancer risk in Sweden
after the Chernobyl Nuclear Power Plant accident in 1986. Uppsala University.
doi:10.33063/diva-544175.

```{r setup}
library(absorbedDose)
```

## Data set

A data set to calculate the doses should look like the generated sample data set <tt>dose_data</tt>. Each person has a unique id. 

```{r data}
print(dose_data, nrow = 10)
```

## Example

The function <tt>calculate_dose()</tt> calls functions for data management 
(<tt>dose_dm()</tt>, <tt>dose_new_variables()</tt>),
for calculation of each dose component (<tt>dose_external()</tt>, <tt>dose_internal_ing()</tt>,
<tt>dose_internal_milk()</tt>, <tt>dose_inhalation()</tt>), and for adding the dose components (<tt>dose_total()</tt>). 
The doses are calculated per raw (per each year of the follow-up time).

Here is an example of calculating the doses to organs Thyroid and Colon.

```{r example}
doses_row <- calculate_dose(dose_data, organ.ls = c("Thyroid", "Colon"))
print(doses_row, nrow = 10)
```

The function <tt>dose_total_per_person()</tt> sums the above calculated absorbed doses for the follow-up period for each person 
(for each id) and for each organ.
```{r example2}
doses_tot <- dose_total_per_person(doses_row)
print(doses_tot, nrow = 10)
```
